Xinhua, Diplomacy, and Chinese Foreign Aid

### 2022-03-10 04_paper/draft1.Rmd
a little more writing

China's motivations for giving foreign aid to developing countries mirror, in many ways, the motivations of Western donors. While recipient needs, in part, predicate Chinese assistance, so do Beijing's political and economic interests. For instance, evidence suggests that China targets more finance toward poorer countries and does not systematically favor authoritarian over democratic recipients [@dreherEtAl2018]. At the same time, China seems to give more aid to countries that import more goods from China, while it gives less aid to those that officially recognize Taiwan [@dreherFuchs2015]. Further, China's Official Financing disproportionately goes to countries with more natural resources, greater UN General Assembly voting alignment, and with more capacity to repay loans [@dreherEtAl2021].

Surveys of policymakers in 126 countries demonstrate the diplomatic impact that these investments have had.^[Reported in blog post ["China's Financial Statecraft: Winning Africa one Yuan at a Time?"](https://www.aiddata.org/blog/chinas-financial-statecraft-winning-africa-one-yuan-at-a-time) by AidData on March 22, 2018. Accessed March 3, 2022]. Compared to a first survey wave in 2014, responses in 2017 showed an increase in evaluations of China's influence with policymakers in developing countries. It would be naive to suppose this is a side-effect, rather than a goal, of China's development finance and other investments.

If its diplomatic motives for giving aid parallel those of other donor governments, visibility is surely a major concern for Beijing. Visual branding of aid has emerged as a growing concern among donors. The United States Agency for International Development (USAID), for example, cites its "new 'brand identity'" as instrumental in promoting the visibility of 2004-2005 tsunami relief.^[See USAID's resource page on branding: [usaid.gov/branding](https://www.usaid.gov/branding).] A US State Department report from 2004 further claims that better branding contributed to a doubling of favorable attitudes toward the US in Indonesia. Pew Research supports this view, finding that nearly 80$\%$ of Indonesians surveyed "said that post-tsunami aid from the US had improved their impression of America."^[["Does humanitarian aid improve America's image?"](https://www.pewresearch.org/global/2012/03/06/does-humanitarian-aid-improve-americas-image/) Published by Pew Research March 6, 2012. Accessed March 3, 2022.]

>>>>

China's motivations for giving foreign aid to developing countries mirror, in many ways, the motivations of Western donors. While recipient needs, in part, predicate Chinese assistance, so do Beijing's political and economic interests. Evidence suggests that China targets more finance toward poorer countries and does not systematically favor authoritarian over democratic recipients [@dreherEtAl2018]. At the same time, China seems to give more aid to countries that import more goods from China, while it gives less aid to those that officially recognize Taiwan [@dreherFuchs2015]. Further, China's Official Financing disproportionately goes to countries with more natural resources, greater UN General Assembly voting alignment, and with more capacity to repay loans [@dreherEtAl2021].

Two waves of surveys conducted by AidData of policymakers in 126 countries demonstrate the diplomatic impact that these investments have had.^[Reported by AidData ["China's Financial Statecraft: Winning Africa one Yuan at a Time?"](https://www.aiddata.org/blog/chinas-financial-statecraft-winning-africa-one-yuan-at-a-time) on March 22, 2018. Accessed March 3, 2022] Compared to a first survey wave in 2014, responses in 2017 showed an increase in evaluations of China's influence with policymakers in developing countries. It would be naive to suppose this is a side-effect, rather than a goal, of China's development finance and other investments.

If its diplomatic motives for giving aid parallel those of other donor governments, visibility is surely a major concern for Beijing. Visual branding of aid has emerged as a growing concern among donors. The United States Agency for International Development (USAID), for example, cited its "new 'brand identity'" as instrumental in promoting the visibility of 2004-2005 tsunami relief to Indonesia.^[See USAID's resource page on branding: [usaid.gov/branding](https://www.usaid.gov/branding).] A US State Department report from 2004 further claims that better branding contributed to a doubling of favorable attitudes toward the US. Pew Research supports this view, finding that nearly 80$\%$ of Indonesians surveyed "said that post-tsunami aid from the US had improved their impression of America."^[["Does humanitarian aid improve America's image?"](https://www.pewresearch.org/global/2012/03/06/does-humanitarian-aid-improve-americas-image/) Published by Pew Research March 6, 2012. Accessed March 3, 2022.]

Given the diplomatic uses of foreign aid, the enduring obscurity that surrounds Beijing's development finance is all the more puzzling. Unlike Western donors, China does not report its financing activities to an international organization or make its activities visible in an online database. To say transparency is out of character for Beijing is an understatement, but foreign aid is one issue area where transparency would arguably serve China's interests. So why the secrecy?

As @fuchsRudyak2019 point out, the obscurity of China's development finance creates the perception among Western donors that China purposefully keeps its aid giving a secret. But as several scholars note, lack of transparency may have more to do with capacity and logistics than intention.^[See @fuchsRudyak2019 for several related citations.] For many decades, a complex and fractured bureaucracy has orchestrated Beijing's development financing. Only in the past few years, as its goals have become ever loftier, did China establish its first true bilateral aid agency---the China International Development Cooperation Agency (CIDCA). However, while established in 2018, much work remains to consolidate the operation and management of China's bilateral aid.^[See this excellent summary entitled ["The Logic Behind China's Foreign Aid Agency"](https://carnegieendowment.org/2019/05/21/logic-behind-china-s-foreign-aid-agency-pub-79154) by the Carnegie Endowment for International Peace published May 21, 2019. Accessed March 10, 2022.]
